---
!---
title: "flat_full.Rmd for working package"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- Run this 'development' chunk -->

<!-- Store every call to library() that you need to explore your functions -->

```{r development, include=FALSE}
library(cluster) 
library(fpc)
library(ggplot2)
library(dplyr)
library(dendextend)
library(UpSetR)
library(nomclust)
library(factoextra)
library(tidyr)
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Data examples

```{=html}
```
```{r development-dataset}
pkgload::load_all(path = here::here(), export_all = FALSE)

datafile <- system.file("data_test.csv", package = "strainselect")
row_data <- read.csv(datafile, sep = ";",header = T)

row_data
```

# prepare_input function

```{=html}
<!--
Include a data table describing bacterial strains according to a number of metadata.
The prepare_input function tests if the strain ID are all different and extracts the metadata selected by the user.
The user can choose to extract the month or the year or both data from the date entered in the initial table.
-->
```
```{r function-prepare_input}
#' prepare_input
#'
#' @param data a dataframe
#' @param col_select a vector
#' @param id a number
#' @param date a number
#' @param m a boolean
#' @param a a boolean
#'
#' @return a dataframe
#' @export
#' @examples
prepare_input <- function (data, col_select=c(1:length(data)), id =1, date=0,
                           m=FALSE, a=TRUE) {
  duplicate <- which(duplicated(data[,id]))
  if (length(duplicate)!=0) {
    stop ( "duplicate ID : " , paste(data[duplicate,id], collapse = ", "))
  } else {
    print("no duplicate ID")
  }
  rownames(data) <- data[,id]
  col_select<-col_select[! col_select %in% date]
  clean_data <- data[,col_select]
  if (date!=0){
    if(a){
      clean_data$ANNEE <- substr(data[,date],7,10)
    }
    if(m){
      clean_data$MOIS <- substr(data[,date],4,5)
    }
  }
  clean_data[is.na(clean_data)] <- "unknown"
  clean_data[] <- lapply(clean_data, factor)
  print(colnames(clean_data))
  str(clean_data)
  clean_data
}
```

```{r examples-prepare_input}

# Example with your dataset in "inst/"
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
# Apply my function
prepare_input(raw_data,col_select = c(7,10,13,14,17,24),
              id=1,date=17,m=FALSE,a=TRUE)
```

```{r tests-prepare_input}
data_test <- matrix("test",nrow = 3,ncol = 5)
as.data.frame(data_test)
test_that("prepare_input works properly and show error if needed", {
  expect_error(prepare_input(data_test))
})
```

# assess_gower function

```{=html}
<!--
Include a data table describing bacterial strains according to a number of metadata.
The assess_gower function runs the prepare_input function on the raw data table.
Gower distances are calculated from the dataframe returned by the prepare_input function. The user can choose to apply weights on the metadata he wants to weight. 
A hierarchical cluster analysis is performed using the gower distances and the "complete" method. 
The dendrogram resulting from the hierarchical cluster analysis is plotted if graph=True.
-->
```
```{r function-assess_gower}
#' assess_gower
#'
#' @param clean_data a dataframe
#' @param weights a vector
#' @param graph a boolean
#'
#' @return a list containing the components clean_data, gower_dist, aggl.clust.c
#' @export
#' @importFrom cluster daisy
#'
#' @examples
assess_gower <- function(clean_data, weights = rep.int(1, p), graph = TRUE){
  p=length(clean_data)
  gower_dist <- daisy(clean_data, metric = "gower", weights = weights)
  aggl.clust.c <- hclust(gower_dist, method = "complete")
  if (graph) {
    plot(aggl.clust.c,main = "Agglomerative, complete linkages")
  }
  return(list(gower_dist=gower_dist, aggl.clust.c=aggl.clust.c))
}
```

```{r examples-assess_gower}
# Example with your dataset in "inst/"
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
prepared_data <- prepare_input(raw_data,col_select = c(7,10,13,14,17,24),
              id=1,date=17,m=FALSE,a=TRUE)
# Apply my function
assess_gower(prepared_data, graph = T)
```

# cstats_table function

```{=html}
<!--
cstat_table create a dataframe including Cluster stats. The avg.silwidth is the data we are most interest in because it will be used to determine the optimal number of cluster to choose if necessary.
The user enters a distance object and a dendogram object resulting from the function assess_gower. He can choose the number of cluster to test and a step to improve the execution time of the function.
-->
```
```{r function-cstats_table}
#' cstats_table
#'
#' @param dist a distance object 
#' @param tree a dendrogram object
#' @param k a number
#' @param l a number
#' @param graph_silhouette a boolean
#'
#' @return a dataframe
#' @export
#'
#' @examples
cstats_table <- function(dist, tree, k,l, graph_silhouette =T) {
   n=k/l
   clust.assess <- c("cluster.number","n","within.cluster.ss","average.within","average.between",
                     "wb.ratio","dunn2","avg.silwidth")
   clust.size <- c("cluster.size")
   stats.names <- c()
   row.clust <- c()
   output.stats <- matrix(ncol = n, nrow = length(clust.assess))
   cluster.sizes <- matrix(ncol =n, nrow = n)
   for(i in 1:n){
       row.clust[i] <- paste("Cluster-", i*l, " size")
   }
    for(i in 1:n){
     print(i*l)
     stats.names[i] <- paste("Test", i*l-1)
     
     for(j in seq_along(clust.assess)){
       output.stats[j, i] <- unlist(cluster.stats(d = dist, clustering = cutree(tree, k = i*l))[clust.assess])[j]
       
     }
     
     for(d in 1:n) {
       cluster.sizes[d, i] <- unlist(cluster.stats(d = dist, clustering = cutree(tree, k = i*l))[clust.size])[d*l]
       dim(cluster.sizes[d, i]) <- c(length(cluster.sizes[i]), 1)
       cluster.sizes[d, i]
     }
   }
   output.stats.df <- data.frame(output.stats)
   cluster.sizes <- data.frame(cluster.sizes)
   cluster.sizes[is.na(cluster.sizes)] <- 0
   rows.all <- c(clust.assess, row.clust)
   # rownames(output.stats.df) <- clust.assess
   output <- rbind(output.stats.df, cluster.sizes)
   colnames(output) <- stats.names[1:n]
   rownames(output) <- rows.all
   is.num <- sapply(output, is.numeric)
   output[is.num] <- lapply(output[is.num], round, 2)
   data_fig <- data.frame(t(output))
   if (graph_silhouette){
     ggplot(data = data_fig, 
            aes(x=cluster.number, y=avg.silwidth)) + 
       geom_point()+
       geom_line()+
       ggtitle("Agglomerative clustering") +
       labs(x = "Num.of clusters", y = "Average silhouette width") +
       theme(plot.title = element_text(hjust = 0.5)) 
   }
 }
```

```{r examples-cstats_table}
# Example with your dataset in "inst/"
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
gower <- assess_gower(prepared_data, graph = T)
# Apply my function
cstats_table(dist = gower$gower_dist, tree = gower$aggl.clust.c, k = 50,l = 10)
```

# select_silhouette function

```{=html}
<!--
After setting the number of cluster chosen, the select_silhouette function selects the strain for each cluster that has the maximum silhouette, i.e. the one that is the best clustered.
If there are multiple strains with a maximum silhouette for the same cluster a random selection is made between the strains concerned.

This function return a data frame with for each strain, its characteristics, its cluster, the corresponding cluster size and the strain selected for its cluster.
-->
```
```{r function-select_silhouette}
#' select_silhouette
#'
#' @param prepared_data a data frame
#' @param dist a distance object
#' @param k a integer
#' @param isdiss logical value specifying wether x is a dissimilarity matrix
#' @param hc_func the hierarchical clustering function to be used
#' @param hc_method he agglomeration method to be used
#'
#' @return a data frame
#' @export
#'
#' @examples
 select_silhouette <- function (prepared_data, dist, k, isdiss = inherits(gower$gower_dist, "dist"),
                               hc_func = "hclust", hc_method = "complete") {
   clust2 <- hcut(dist, k, isdiss,hc_func, hc_method)
   silh <- fviz_silhouette(clust2)
   selected_data <- prepared_data
   selected_data$cluster <- clust2$cluster
   n=nrow(selected_data)
   selection <- rep(NA,k)
   for(i in 1:k){
     max_silh_i <- silh$data[which(silh$data$sil_width==max(silh$data$sil_width[silh$data$cluster==i])&silh$data$cluster==i),]
     if (nrow(max_silh_i)==1){
       selection[i] <- as.character(max_silh_i$name)
     } else {
       selection[i] <- as.character(sample(max_silh_i$name,1))
     } 
   }
   for (i in 1:n){
     for (j in 1:k){
       if (selected_data$cluster[i]==j){
         selected_data$size[i] <- clust2$size[j]
         selected_data$selected[i] <- selection[j]
         
       }
     }
   }
   selected_data
 }
```

```{r examples-select_silhouette}
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
gower <- assess_gower(prepared_data, graph = T)
prepared_data <- prepare_input(raw_data,col_select = c(7,10,13,14,17,24),
              id=1,date=17,m=FALSE,a=TRUE)

 select_silhouette(dist=gower$gower_dist,prepared_data = prepared_data, k=50, isdiss = inherits(gower$gower_dist, "dist"),hc_func = "hclust", hc_method = "complete")
```

# prepare_upset function

```{=html}
<!--
The prepare_upset allows to prepare the dataframe used to visualize the interactions. it uses the modalities of the different variables of prepared_data dataframe as columns and fills the rows with 1 if the strain is concerned by the modality and 0 otherwise.
-->
```
```{r function-prepare_upset}
#' prepare_upset
#'
#' @param prepared_data a dataframe
#'
#' @return a dataframe
#' @export
#'
#' @examples
prepare_upset<- function(prepared_data){
 upset_data <- data.frame(row.names = rownames(prepared_data))
 metadata <- colnames(prepared_data)
 for (i in 1:ncol(prepared_data)){
   vect <- unique(prepared_data[,i])
   tab <- matrix(0,nrow(prepared_data),length(vect))
   col.names <- NULL
   for (j in 1:length(vect)){
     col.names <- c(col.names,paste(metadata[i],"_",vect[j]))
     for (k in 1:nrow(prepared_data)){
       if (is.na(prepared_data[k,i])==F){
         if(prepared_data[k,i]==vect[j]){
           tab[k,j] <- 1
         }
       }
     }
   }
   colnames(tab) <- col.names
   upset_data <- cbind(upset_data,tab)
 } 
 upset_data
}

```

```{r examples-prepare_upset}
# Example with your dataset in "inst/"
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
prepared_data <- prepare_input(raw_data,col_select = c(7,10,13,14,17,24),
              id=1,date=17,m=FALSE,a=TRUE)
# Apply my function
prepare_upset(prepared_data)
```

# define_profile function

```{=html}
<!--
This function uses the output of pepare_upset function to create a new data frame with a column "profiles" resulting  from the concatenation of the columns of the input data frame and a colonne that identify the group of the strain. 
Strain are grouped by identical profiles which allows to consider also the strains with a unique profile.
-->
```
```{r function-define_profiles}
#' define_profiles
#'
#' @param data a data frame
#' @param sep  Separator to use between values
#'
#' @return a data frame
#' @export
#'
#' @examples
define_profiles <- function (data, sep = ""){
  profiles <- unite(data,profiles,c(1:ncol(upset_data)), sep = sep)
  profiles_freq <- as.data.frame(table(profiles))
  profiles_freq$group <- c(1:nrow(profiles_freq))
  n=nrow(profiles)
  for (i in 1:n){
    for (j in 1:nrow(profiles_freq)){
      if (profiles$profiles[i]==profiles_freq$profiles[j]){
        profiles$group[i]<-profiles_freq$group[j]
      }
    }
  }
  profiles$names<-rownames(profiles)
  
  profiles
} 
```

```{r examples-define_profile}
# Example with your dataset in "inst/"
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
prepared_data <- prepare_input(raw_data,col_select = c(7,10,13,14,17,24),
              id=1,date=17,m=FALSE,a=TRUE)
upset_data <- prepare_upset(prepared_data)
# Apply my function
define_profiles(upset_data)
```

# select_profiles function

```{=html}
<!--
Because the define_profiles function groups strains by identical profile, select_profiles function randomly choose one strain of each group. 
-->
```
```{r function-select_profiles}
#' select_profiles
#'
#' @param profiles a data frame
#' @param prepared_data a data frame
#'
#' @return a data frame
#' @export
#'
#' @examples
select_profiles <- function(profiles, prepared_data){
  selected_data <- prepared_data
  selected_data$group <- profiles$group
  selected_data$size <- profiles$freq
  k<- max(profiles$group)
  n=nrow(selected_data)
  selection <-rep(NA,k)
  for (j in 1:k){
      selection[j] <-sample(profiles$names[profiles$group==j],1)
  }
  for (i in 1:n){
    for (j in 1:k){
      if (selected_data$group[i]==j){
        selected_data$selected[i] <- selection[j]
      }
    }
  }
  selected_data
}
```

```{r examples-select_profile}
# Example with your dataset in "inst/"
datafile <- system.file("data_test.csv", package = "strainselect")
raw_data <- read.csv(datafile, sep = ";",header = T)
prepared_data <- prepare_input(raw_data,col_select = c(7,10,13,14,17,24),
              id=1,date=17,m=FALSE,a=TRUE)
upset_data <- prepare_upset(prepared_data)
upset_profiles <- define_profiles(upset_data)
# Apply my function
select_profiles(upset_profiles,prepared_data)
```

```{r development-inflate, eval=FALSE}
# Keep eval=FALSE to avoid infinite loop in case you hit the knit button
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_full.Rmd", vignette_name = "Get started")
```

# Inflate your package

You're one inflate from paper to box. Build your package from this very Rmd using `fusen::inflate()`

-   Verify your `"DESCRIPTION"` file has been updated
-   Verify your function is in `"R/"` directory
-   Verify your test is in `"tests/testthat/"` directory
-   Verify this Rmd appears in `"vignettes/"` directory
